Social Attention for Autonomous Decision-Making in Dense Traffic

### 2019-09-06 main.tex
Update

Automated Decision Making
Modular pipeline: perception, high level representation, decision: behaviour planning, trajectory, control
large litterature on perception
trajectory optimization and control is well understood

behaviour planning by rule-based systems, fsm
but behaviour planning is difficult, complex scenes, in particular interactions with other drivers difficult to model explicitly.
 -> leverage (deep) learning method
Learning from demonstration: nice results for lane keeping, not so much for interaction, also require costly demonstration .

Conversely, leverage simulation:
Reinforcement learning:
Framing as an MDP, choose a good representation.
List of vehicles with their coordinates 
but variable size, + combinatorial representation
Occupancy grid / images : fixed size, tradeoff:  low accuracy vs exploding size

@@

In the state:
Kinematics
Not fixed size
Occupancy grid

In the model:

Weight sharing ?
Socially aware}

@@

\section{Experiments}
\subsection{Description of the environment}

@@

\begin{table}
	\centering
	\begin{tabular}{lcc}
		\toprule
		Architecture & MLP & Ego-Attention \\
		\midrule 
		Layers sizes & [128, 128] & [64, 64], 64, [64, 64] \\
		Number of parameters & \textbf{3.0e4} & 3.4e4 \\
		Training Time (w.r.t. MLP) & \textbf{1} & 1.25 \\
		Mean total reward & 3.6 & \textbf{6.4} \\ 
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}
	\centering
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/total_rewards}
		\caption{Cumulative reward}
	\end{subfigure}
\\
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/total_length}
		\caption{Episode length}
	\end{subfigure}
\\
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/collision}
		\caption{Probability of collision}
	\end{subfigure}
\end{figure}

>>>>

In the last decades, the problem of \emph{behavioural planning} -- that is, high-level decision-making in the context of autonomous driving -- has arguably received less attention and seen less progress than the other components of the typical robotics pipeline: perception and control \citep{Gonzalez2016}. Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}. As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviour is uncertain and difficult to model explicitly.

This observation has led the community to turn to learning-based methods, which bear the promise of leveraging data to automatically learn a complex driving policy. In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Bansal2018}. Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep{Cardamone2009, Ross2011}.

Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process (MDP) plays an important part in the design of the system. In particular, the choice of state and action spaces can have a significant influence on the performance of resulting policy. Describing a driving scene typically involves choosing a representation the road and vehicles. The simplest form: a list of vehicles with their coordinates, orientations, velocities. But variable size and order-dependent. Another common format is a spatial matrix, either Cartesian (images, occupancy grid) or polar (distances in angular sectors). Fixed size and order independent, but accuracy/size tradeoff. Challenge: keep a small state size, but enforce that the model is invariant to permutations.

@@

A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity. The composite state (or joint-state) of a road traffic with one ego-vehicle (denoted $s_0$) and N other vehicles can then be described by the set of the states of all vehicles.

\begin{equation*}
s = \left\lbrace s_k \right\rbrace_{k \in [0, N]}\qquad
\text{where}\qquad
s_k = \begin{bmatrix}
x_k & y_k & v^x_k & v^y_k & \cos\psi_k & \sin \psi_k
\end{bmatrix}^T
\end{equation*}

\begin{figure}[tp]
	\centering
	\includegraphics[width=0.25\textwidth]{img/coordinates}
	\includegraphics[width=0.25\textwidth]{img/map}
	\caption{The coordinates (left) and spatial-grid (right) representations}
\end{figure}

This representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.

This encoding is efficient in the sense that it uses the smallest quantity of information necessary to represent the scene. However, it lacks two important properties. First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs, but also for concerns of varying computational complexity. Second, we expect a driving policy to be \emph{permutation invariant}, i.e. not to be dependent on the order in which other traffic participants are listed. Ideally, this property should be enforced and not relying on the coverage of the $N!$ possible permutations of any given traffic state. Formally, we require that:
\begin{equation*}
\pi(\cdot|(s_1,\dotsc,s_N)) = \pi(\cdot|(s_{\sigma(1)},\dotsc,s_{\sigma(N)})) \quad\quad\quad \forall\sigma \in \mathfrak{S}_N
\end{equation*}

A popular way to address this limitations is to use a spatial matrix representation. Instead of explicitly representing spatial information as variables $x, y$ within the state $\{s_k=(x_k,y_k,f_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_ij$ organized in a tensor, where the $i,j$ indexes refer to cells of a quantization of the 2D-space. Note that the size of this tensor is related to the area covered divided by the length of the quantization, which denotes a trade-off between accuracy and computational burden. 
In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep{Isele2017, Fridman2018, Bansal2018, Rehder2017c}. Another example is the use of top-view RGB images \citep{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.

Transition:
This permutation invariance property can be implemented within the policy architecture, as done in \citep{Chen2017} or \citep{Qi2016}, but also directly in the state representation.
Weight sharing and reduce operator that is invariant to permutation eg max in \citep{Chen2017} and sum in \citep{point_net}.
A particular instantiation of this idea: attention.

@@

\section{Experiments}
\paragraph{The environment}

@@

\begin{table}
	\centering
	\begin{threeparttable}
    \caption{Results}
	\begin{tabular}{lccc}
		\toprule
		Architecture & MLP Coordinates & MLP Grid & Ego-Attention \\
		\midrule 
		Layers sizes & [128, 128] &  [64, 64] & [64, 64], 64, [64, 64] \\
		Number of parameters & \textbf{3.0e4} & ? & 3.4e4 \\
		Mean total reward & 3.7 & 4.5 & \textbf{6.4} \\ 
		\bottomrule
	\end{tabular}
%	\begin{tablenotes}
%		\small
%		\item 
%	\end{tablenotes}
	\end{threeparttable}
\end{table}

\begin{figure}
	\centering
	\begin{subfigure}[t]{.49\linewidth}
		\centering\includegraphics[width=\linewidth]{img/total_reward}
	\end{subfigure}
	\begin{subfigure}[t]{.49\linewidth}
		\centering\includegraphics[width=\linewidth]{img/length}
	\end{subfigure}
	\begin{subfigure}[t]{.49\linewidth}
		\centering\includegraphics[width=\linewidth]{img/velocity}
	\end{subfigure}
	\begin{subfigure}[t]{.49\linewidth}
	\centering\includegraphics[width=\linewidth]{img/distance}
    \end{subfigure}
\end{figure}
